---
title: "FIMO Results Analysis for ABC Predicted Enhancers"
output: html_document
---

```{r setup, include = F}
knitr::opts_chunk$set(message = F, warning = F)
```

```{r libaries, include = F}
library(data.table)
library(patchwork)
source('parseGff.R')
source('tfCoverage.R')
```

```{r make_controls, include = F}
# make control bed files for comparison of coverage
# 10 random upstream 500bp regions 
# starts <- sample(seq(from = 117404938, to = 117470238, by = 500), 10, replace=F)
# contBed <- data.table(chr = rep("chr7", 10),
#                       start = starts,
#                       end = starts + 500,
#                       name = rep("control", 10))
# fwrite(contBed, 'controlRegs.bed', quote = F, row.names = F, col.names = F, sep = '\t')

```

```{r load_fimo, include = F}
# results with Hv13
# JASPAR results were noisy and redundant
bsHv <- as.data.table(gff2df('data/FIMO/basalHv13.gff', hasAlias = F))
clHv <- as.data.table(gff2df('data/FIMO/ciliatedHv13.gff', hasAlias = F))
scHv <- as.data.table(gff2df('data/FIMO/secretoryHv13.gff', hasAlias = F))
contHv <- as.data.table(gff2df('data/FIMO/controlRegs.gff', hasAlias = F))
```

```{r load_predictions, include = F}
# col names
abcCols <- c("chr", "start", "end", "name",
             "class", "activity_base", "activity_base_enh",
             "activity_base_squared_enh", "normalized_atac_enh",
             "enh_midpoint", "enh_idx", "TargetGene", "TargetGeneTSS",
             "TargetGeneExpression", "TargetGenePromoterActivityQuantile",
             "TargetGeneIsExpressed", "TargetGeneEnsembl_ID", "normalized_atac_prom",
             "gene_idx", "distance", "isSelfPromoter", "powerlaw_contact", 
             "powerlaw_contact_reference", "ABC.Score.Numerator", "ABC.Score",
             "powerlaw.Score.Numerator", "powerlaw.Score", "CellType")

# predictions by cell type
bsPred <- fread('data/predictions/basalPredictions.tsv', col.names = abcCols)
clPred <- fread('data/predictions/ciliatedPredictions.tsv', col.names = abcCols)
scPred <- fread('data/predictions/secretoryPredictions.tsv', col.names = abcCols)

# load control bed
contBed <- fread('controlRegs.bed', col.names = c('chr', 'start', 'end', 'name'))
```

## Plotting Coverage for Predicted Enhancers

Here we compare FIMO results in different enhancer regions and quantify overall binding site potential by computing "coverage", i.e. what % of the region is covered by a predicted binding site.

```{r plot_example}
# same as chunk above but with function
multiCovplot <- function(predictions, fimores) {
  plts <- lapply(1:nrow(predictions), function(x) gffCov(fimores, qCut = 0.05,
                                                         bSize = 20, startLim = predictions[x]$start,
                                                         endLim = predictions[x]$end,
                                                         title = predictions[x]$name))
  return(wrap_plots(plts))
} 

multiCovplot(bsPred, bsHv)
```

## Control Plot
Coverage for 10 randomly selected 500bp regions in the upstream CFTR region (between promoter and CTCF site TAD boundary)
```{r control_plot}
multiCovplot(contBed, contHv)
```

```{r make_plots, include = F}
ggsave('plots/basalHvCovplot.png', multiCovplot(bsPred, bsHv))
ggsave('plots/ciliatedHvCovplot.png', multiCovplot(clPred, clHv))
ggsave('plots/secretoryHvCovplot.png', multiCovplot(scPred, scHv))
ggsave('plots/controlRegsCovplot.png', multiCovplot(contBed, contHv))
```

```{r cov_tables, include = F}
# add coverage column to samples
bsPred[, coverage := mapply(getOverlap, start, end, 
                            MoreArgs=list(mergeRanges(bsHv[qvalue < 0.05])))]
clPred[, coverage := mapply(getOverlap, start, end, 
                            MoreArgs=list(mergeRanges(clHv[qvalue < 0.05])))]
scPred[, coverage := mapply(getOverlap, start, end, 
                            MoreArgs=list(mergeRanges(scHv[qvalue < 0.05])))]
contBed[, coverage := mapply(getOverlap, start, end,
                             MoreArgs=list(mergeRanges(contHv[qvalue < 0.05])))]
```

## TF Motifs Found in Specific Enhancer Regions

Here we look at the FIMO results for each enhancer region. The example below shows the number of sites as well as the individual transcription factors with predicted sites given a q-value threshold of 0.05. The region used as an example here is the -44kb upstream enhancer site found in the basal cell scATAC data.

```{r inspect_motifs}
# get unique tfs from given range
s <- 117479505
e <- 117480212
fltr <- bsHv[start < e & end > s & qvalue < 0.05]
nrow(fltr)
unique(fltr$Alias)
```

# Variations in q-value vs p-value, FIMO score

```{r qval_testing}
# p val vs q val
ggplot(bsHv, aes(x=-log10(pvalue), y=-log10(qvalue))) + geom_point() + ggtitle('p-val vs q-val for FIMO results')
ggplot(bsHv, aes(x=score, y=-log10(qvalue))) + geom_point() + ggtitle('score vs q-val for FIMO results')
ggplot(bsHv, aes(x=score, y=-log10(pvalue))) + geom_point() + ggtitle('score vs p-val for FIMO results')
```

```{r length_norm}
bsHv[, length := end - start]
ggplot(bsHv, aes(x=-log10(pvalue), y=-log10(qvalue), color=length)) + geom_point() + ggtitle('p-val vs q-val for FIMO results')

head(bsHv[order(length, decreasing = T), .(Alias, pvalue, qvalue, sequence)], 20)
```